---
title: |
  STA 336: Statistical Machine Learning
subtitle: |
  NFL Wide Receiver Contract Projection
author: |
  Rento Saijo, Atticus Roop, and Manahil Shoaib
date: 'May 1, 2026'
fontsize: 11pt
documentclass: article
geometry: top=1in, bottom=1in, left=1in, right=1in
urlcolor: blue
linkcolor: blue
citecolor: blue
bibliography: paper.bib
link-citations: true
output:
  bookdown::pdf_document2:
    keep_tex: true
    toc: false
    fig_width: 6.5
    fig_height: 4
    fig_caption: true

header-includes:
  - |
    \usepackage{xcolor}
    \usepackage{hyperref}
    \hypersetup{colorlinks=true, linkcolor=blue, urlcolor=blue, citecolor=blue}
    \usepackage{float}
    \usepackage{fvextra}
    \usepackage{fancyhdr}
    \usepackage{lastpage}
    \usepackage{soul}
    \usepackage{etoolbox}
    \usepackage{microtype}
    \usepackage{tcolorbox}
    \usepackage{enumitem}
    \usepackage{fancyvrb}
    \usepackage{xspace}
    \allowdisplaybreaks
    \setcounter{tocdepth}{2}
    \setcounter{secnumdepth}{0}
    \floatplacement{figure}{H}
    \makeatletter
    \newcommand{\@subtitle}{}
    \newcommand{\subtitle}[1]{\gdef\@subtitle{#1}}
    \newcommand{\DocSubtitle}{\@subtitle}
    \renewcommand{\maketitle}{
      \thispagestyle{plain}
      \null
      \vfill
      \begin{center}
        {\Large \textsc{\@title} \par}\vskip 0.5em
        {\LARGE \bfseries \@subtitle \par}\vskip 0.75em
        {\large \@author \par}\vskip 0.5em
        {\normalsize \@date \par}
      \end{center}
      \vfill
      \clearpage
    }
    \AfterEndEnvironment{Highlighting}{\setcounter{CodeLine}{\numexpr\value{CodeLine}+\FV@CodeLineNo\relax}}
    \makeatother
    \fancypagestyle{plain}{
      \fancyhf{}
      \renewcommand{\headrulewidth}{0pt}
      \renewcommand{\footrulewidth}{0pt}
    }
    \pagestyle{fancy}
    \fancyhf{}
    \setlength{\headheight}{14pt}
    \fancyfoot[C]{\small\textsc{Page}~\thepage~\textsc{of}~\pageref{LastPage}}
    \fancyhead[L]{\textsc{Saijo, Roop, and Shoaib}}
    \fancyhead[R]{\textsc{NFL WR Contract Projection}}
    \fancyhead[C]{}
    \renewcommand{\headrulewidth}{0.5pt}
    \renewcommand{\footrulewidth}{0.5pt}
    \newcounter{CodeLine}
    \newcounter{cell}
    \AtBeginEnvironment{Highlighting}{\stepcounter{cell}}
    \DefineVerbatimEnvironment{Highlighting}{Verbatim}{
      frame=single,
      rulecolor=\color{black},
      framesep=2mm,
      label=\footnotesize\textbf{Cell \thecell},
      labelposition=topline,
      commandchars=\\\{\},
      breaklines=true,
      breakanywhere=false,
      breaksymbol={},
      numbers=left,
      numbersep=3pt,
      firstnumber=\numexpr\value{CodeLine}+1\relax,
      fontsize=\small
    }
    \DefineVerbatimEnvironment{verbatim}{Verbatim}{
      breaklines=true,
      breakanywhere=true,
      numbers=left,
      numbersep=3pt,
      firstnumber=1,
      fontsize=\footnotesize
    }
    \DefineVerbatimEnvironment{snippet}{Verbatim}{
      breaklines=true,
      breakanywhere=true,
      fontsize=\footnotesize,
      frame=single,
      listparameters=\setlength{\topsep}{\baselineskip}\setlength{\partopsep}{0pt}
    }
    \newtcolorbox{problem}[1][]{
      title={\textsc{#1}}
    }
    \newenvironment{alphenum}{
      \begin{enumerate}[
        label=(\alph*),
        itemsep=3pt,
        parsep=0pt,
        topsep=6pt
      ]
    }{
      \end{enumerate}
    }
    \newenvironment{items}{
      \begin{itemize}[
        itemsep=3pt,
        parsep=0pt,
        topsep=6pt
      ]
    }{
      \end{itemize}
    }
    \renewcommand{\contentsname}{Table of Contents}
    \newcommand{\E}{\mathrm{E}}
    \newcommand{\Var}{\mathrm{Var}}
    \newcommand{\R}{\textsf{R}\xspace}
    \newcommand{\SD}{\mathrm{SD}}
    \newcommand{\SE}{\mathrm{SE}}
---

```{r setup, include = FALSE}
# ----- Setup ----- #

knitr::opts_chunk$set(
  echo       = FALSE,
  warning    = FALSE,
  message    = FALSE,
  fig.align  = 'center',
  fig.pos    = 'H',
  out.width  = '\\linewidth'
)
options(width = 125)
options(scipen = 999)

suppressMessages(library(dplyr))
suppressMessages(library(forcats))
suppressMessages(library(ggplot2))
suppressMessages(library(knitr))
suppressMessages(library(kableExtra))
suppressMessages(library(readr))
suppressMessages(library(scales))
suppressMessages(library(stringr))
suppressMessages(library(tibble))
suppressMessages(library(tidyr))

font_files <- c(
  regular    = '/Users/rsai_91/Library/Fonts/cmu.serif-roman.ttf',
  bold       = '/Users/rsai_91/Library/Fonts/cmu.serif-bold.ttf',
  italic     = '/Users/rsai_91/Library/Fonts/cmu.serif-italic.ttf',
  bolditalic = '/Users/rsai_91/Library/Fonts/cmu.serif-bolditalic.ttf'
)
if (requireNamespace('sysfonts', quietly = TRUE) &&
    requireNamespace('showtext', quietly = TRUE) &&
    all(file.exists(font_files))) {
  sysfonts::font_add(
    family     = 'cmuserif',
    regular    = font_files[['regular']],
    bold       = font_files[['bold']],
    italic     = font_files[['italic']],
    bolditalic = font_files[['bolditalic']]
  )
  showtext::showtext_opts(dpi = 500)
  showtext::showtext_auto(enable = TRUE)
  knitr::opts_chunk$set(dev = 'CairoPDF')
  base_family <- 'cmuserif'
} else {
  base_family <- 'serif'
}

theme_set(
  theme_minimal(base_family = base_family, base_size = 10) +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.line = element_line(linewidth = 0.25, color = 'grey35')
    )
)

# ----- Helpers ----- #

table_latex <- function(
  data,
  col_names  = NULL,
  caption    = NULL,
  escape     = TRUE,
  align      = NULL,
  wrap_col   = NULL,
  wrap_width = NULL,
  font_size  = 8
) {
  if (is.null(align)) {
    align <- rep('c', ncol(data))
  }
  table_out <- knitr::kable(
    data,
    format    = 'latex',
    booktabs  = TRUE,
    escape    = escape,
    col.names = col_names,
    caption   = caption,
    linesep   = '',
    align     = align
  )
  table_out <- kableExtra::kable_styling(
    table_out,
    position      = 'center',
    latex_options = c('HOLD_position'),
    font_size     = font_size
  )
  if (!is.null(wrap_col) && !is.null(wrap_width)) {
    table_out <- kableExtra::column_spec(table_out, wrap_col, width = wrap_width)
  }
  table_out
}

comma_num <- function(x, digits = 0) {
  format(round(x, digits), big.mark = ',', scientific = FALSE, trim = TRUE, nsmall = digits)
}

pct_num <- function(x, digits = 1) {
  paste0(format(round(100 * x, digits), nsmall = digits), '%')
}

money_m <- function(x, digits = 2) {
  paste0('$', format(round(x, digits), nsmall = digits), 'M')
}

clean_feature <- function(x) {
  x |>
    str_replace_all('_', ' ') |>
    str_replace_all('([a-z])([A-Z])', '\\1 \\2') |>
    str_replace_all('Prev1', 'Previous 1 ') |>
    str_replace_all('Prev2', 'Previous 2 ') |>
    str_replace_all('Weighted', 'Weighted ') |>
    str_squish()
}

# ----- Data ----- #

train_data    <- read_csv('data/train.csv', show_col_types = FALSE)
validate_data <- read_csv('data/validate.csv', show_col_types = FALSE)
test_data     <- read_csv('data/test.csv', show_col_types = FALSE)

term_baseline <- read_csv('results/term_baseline.csv', show_col_types = FALSE)
aav_baseline  <- read_csv('results/aavPerc_basline.csv', show_col_types = FALSE)
term_opt      <- read_csv('results/term_optimized.csv', show_col_types = FALSE)
aav_opt       <- read_csv('results/aavPerc_optimized.csv', show_col_types = FALSE)

summary_xgb   <- read_csv('final/summary_xgb.csv', show_col_types = FALSE)
summary_glm   <- read_csv('final/summary_glm.csv', show_col_types = FALSE)
scenarios_xgb <- read_csv('final/scenarios_xgb.csv', show_col_types = FALSE)
scenarios_glm <- read_csv('final/scenarios_glm.csv', show_col_types = FALSE)

term_xgb_imp  <- read_csv('models/term_xgb_final_importance.csv', show_col_types = FALSE)
aav_xgb_imp   <- read_csv('models/aavPerc_xgb_final_importance.csv', show_col_types = FALSE)
term_ll_coef  <- read_csv('models/term_ll_final_coefficients.csv', show_col_types = FALSE)
aav_rl_coef   <- read_csv('models/aavPerc_rl_final_coefficients.csv', show_col_types = FALSE)

term_xgb_model <- readRDS('models/term_xgb_final.rds')
aav_xgb_model  <- readRDS('models/aavPerc_xgb_final.rds')

split_tbl <- tibble(
  split = c('Train', 'Validate', 'Test'),
  role = c(
    'Historical signed contracts used for fitting and tuning',
    'Signed 2026 contracts used for held-out model comparison',
    'Unsigned 2026 free-agent candidates scored by the final models'
  ),
  rows = c(nrow(train_data), nrow(validate_data), nrow(test_data)),
  players = c(
    n_distinct(train_data$playerName),
    n_distinct(validate_data$playerName),
    n_distinct(test_data$playerName)
  ),
  start_years = c(
    paste0(min(train_data$startYear), '-', max(train_data$startYear)),
    as.character(unique(validate_data$startYear)),
    as.character(unique(test_data$startYear))
  )
)

term_dist_tbl <- bind_rows(
  train_data |> mutate(split = 'Train'),
  validate_data |> mutate(split = 'Validate')
) |>
  count(split, term) |>
  group_by(split) |>
  mutate(pct = n / sum(n)) |>
  ungroup()

coverage_tbl <- tibble(
  family = c('Observed ESPN signing date', 'NFL combine', 'Regular receiving stats', 'Snap usage', 'Next Gen receiving', 'Team context', 'Market context'),
  train_coverage = c(
    mean(!is.na(train_data$dateOfSigningObserved)),
    mean(train_data$combineAvailable %in% TRUE, na.rm = TRUE),
    mean(train_data$regPrev1Available | train_data$regPrev2Available, na.rm = TRUE),
    mean(train_data$snapPrev1Available | train_data$snapPrev2Available, na.rm = TRUE),
    mean(train_data$ngsPrev1Available | train_data$ngsPrev2Available, na.rm = TRUE),
    mean(train_data$teamPrev1Available | train_data$teamPrev2Available, na.rm = TRUE),
    mean(train_data$marketPrev1Available | train_data$marketPrev2Available, na.rm = TRUE)
  )
)

opt_tbl <- bind_rows(
  term_opt |>
    transmute(
      task = 'Term',
      model = dplyr::case_when(
        modelId == 'term_xgb' ~ 'XGBoost',
        modelId == 'term_rf'  ~ 'Random forest',
        modelId == 'term_ll'  ~ 'Lasso multinomial',
        TRUE ~ modelId
      ),
      metric_1 = sprintf('Log loss = %.3f', mnLogLoss),
      metric_2 = sprintf('Brier = %.3f', brier),
      metric_3 = sprintf('Accuracy = %.1f%%', 100 * accuracy),
      cv_metric = sprintf('CV log loss = %.3f', cvMetricMean)
    ),
  aav_opt |>
    transmute(
      task = 'AAV%',
      model = dplyr::case_when(
        modelId == 'aavPerc_xgb' ~ 'XGBoost',
        modelId == 'aavPerc_rf'  ~ 'Random forest',
        modelId == 'aavPerc_rl'  ~ 'Ridge linear',
        TRUE ~ modelId
      ),
      metric_1 = sprintf('RMSE = %.4f', rmse),
      metric_2 = sprintf('MAE = %.4f', mae),
      metric_3 = sprintf('R-squared = %.3f', rsq),
      cv_metric = sprintf('CV RMSE = %.4f', cvMetricMean)
    )
)

top_projection_tbl <- summary_xgb |>
  arrange(desc(expectedAavMillions)) |>
  transmute(
    Player = playerName,
    Age = ageAtSigning,
    `Previous Team` = prevSignedTeam,
    `Expected Term` = round(expectedTerm, 2),
    `P(1 Year)` = pct_num(.pred_1, 1),
    `P(2+ Years)` = pct_num(.pred_2 + .pred_3 + .pred_4 + .pred_5, 1),
    `Expected AAV` = money_m(expectedAavMillions, 2),
    `Modal AAV` = money_m(modalAavMillions, 2)
  ) |>
  slice_head(n = 10)

main_projection_tbl <- top_projection_tbl |>
  slice_head(n = 6)

compact_projection_tbl <- summary_xgb |>
  arrange(desc(expectedAavMillions)) |>
  transmute(
    Player = playerName,
    `Expected Term` = round(expectedTerm, 2),
    `P(1 Year)` = pct_num(.pred_1, 1),
    `P(2+ Years)` = pct_num(.pred_2 + .pred_3 + .pred_4 + .pred_5, 1),
    `Expected AAV` = money_m(expectedAavMillions, 2)
  ) |>
  slice_head(n = 5)

main_summary_tbl <- tibble(
  Component = c(
    'Training data',
    'Validation data',
    'Scored 2026 market',
    'Best term model',
    'Best compensation model',
    'Market-level forecast'
  ),
  Evidence = c(
    paste0(comma_num(nrow(train_data)), ' signed veteran WR contracts from ', min(train_data$startYear), '-', max(train_data$startYear)),
    paste0(nrow(validate_data), ' signed 2026 contracts; ', pct_num(mean(validate_data$term == 1), 1), ' were one-year deals'),
    paste0(nrow(test_data), ' unsigned 2026 WR free agents'),
    'XGBoost with log loss 0.537, Brier 0.057, accuracy 82.4%',
    'XGBoost with RMSE 0.0057, MAE 0.0026, R-squared 0.903',
    paste0('Mean one-year probability ', pct_num(mean(summary_xgb$.pred_1), 1), '; median expected AAV ', money_m(median(summary_xgb$expectedAavMillions), 2))
  )
)

baseline_lookup <- bind_rows(
  term_baseline |>
    transmute(
      task = 'Term',
      model_id = modelId,
      log_loss = mnLogLoss,
      brier = brier,
      accuracy = accuracy,
      rmse = NA_real_,
      mae = NA_real_,
      rsq = NA_real_
    ),
  aav_baseline |>
    transmute(
      task = 'AAV%',
      model_id = modelId,
      log_loss = NA_real_,
      brier = NA_real_,
      accuracy = NA_real_,
      rmse = rmse,
      mae = mae,
      rsq = rsq
    )
)

baseline_screen_tbl <- tibble(
  Task = c('Term', 'Term', 'AAV%', 'AAV%'),
  Role = c('Predictive', 'Interpretable', 'Predictive', 'Interpretable'),
  `Best Untuned Baseline` = c('XGBoost', 'Lasso', 'XGBoost', 'Ridge'),
  `Validation Evidence` = c(
    sprintf(
      'Log loss %.3f; Brier %.3f; acc. %.1f%%',
      baseline_lookup$log_loss[baseline_lookup$task == 'Term' & baseline_lookup$model_id == 'xgboost'],
      baseline_lookup$brier[baseline_lookup$task == 'Term' & baseline_lookup$model_id == 'xgboost'],
      100 * baseline_lookup$accuracy[baseline_lookup$task == 'Term' & baseline_lookup$model_id == 'xgboost']
    ),
    sprintf(
      'Log loss %.3f; Brier %.3f; acc. %.1f%%',
      baseline_lookup$log_loss[baseline_lookup$task == 'Term' & baseline_lookup$model_id == 'lasso_glmnet'],
      baseline_lookup$brier[baseline_lookup$task == 'Term' & baseline_lookup$model_id == 'lasso_glmnet'],
      100 * baseline_lookup$accuracy[baseline_lookup$task == 'Term' & baseline_lookup$model_id == 'lasso_glmnet']
    ),
    sprintf(
      'RMSE %.4f; MAE %.4f; R-squared %.3f',
      baseline_lookup$rmse[baseline_lookup$task == 'AAV%' & baseline_lookup$model_id == 'xgboost'],
      baseline_lookup$mae[baseline_lookup$task == 'AAV%' & baseline_lookup$model_id == 'xgboost'],
      baseline_lookup$rsq[baseline_lookup$task == 'AAV%' & baseline_lookup$model_id == 'xgboost']
    ),
    sprintf(
      'RMSE %.4f; MAE %.4f; R-squared %.3f',
      baseline_lookup$rmse[baseline_lookup$task == 'AAV%' & baseline_lookup$model_id == 'ridge_glmnet'],
      baseline_lookup$mae[baseline_lookup$task == 'AAV%' & baseline_lookup$model_id == 'ridge_glmnet'],
      baseline_lookup$rsq[baseline_lookup$task == 'AAV%' & baseline_lookup$model_id == 'ridge_glmnet']
    )
  )
)

xgb_summary_stats <- summary_xgb |>
  summarise(
    mean_expected_term = mean(expectedTerm),
    median_expected_term = median(expectedTerm),
    max_expected_term = max(expectedTerm),
    mean_expected_aav = mean(expectedAavMillions),
    median_expected_aav = median(expectedAavMillions),
    max_expected_aav = max(expectedAavMillions),
    mean_p1 = mean(.pred_1),
    mean_p3plus = mean(.pred_3 + .pred_4 + .pred_5)
  )

dashboard_url <- 'https://rentosaijo.github.io/NFLxC/'
```

# Abstract

This project studies whether public data can support useful projections for the 2026 veteran wide receiver contract market. The central questions are whether contract term can be predicted, whether annual average value (AAV) can be predicted as a percentage of the NFL salary cap, and which player, contract, production, and market factors appear most important in those predictions. The analysis uses Spotrac wide receiver contract exports as the contract spine and enriches those records with nflverse player, roster, combine, production, snap, Next Gen Stats, team, and schedule data, plus ESPN transaction dates when a likely signing match is available [@spotrac; @nflreadr; @espntransactions]. The modeling design separates contract term from annual price because the two outcomes represent different market decisions. Term measures commitment, while cap-share AAV measures price conditional on that commitment. The strongest predictive pair uses XGBoost for both tasks [@xgboost], while lasso multinomial logistic regression and ridge regression provide interpretable audit models [@glmnet]. On the signed 2026 validation set, the selected term model achieved log loss of 0.537, Brier score of 0.057, and accuracy of 82.4%; the selected compensation model achieved RMSE of 0.0057, MAE of 0.0026, and \(R^2\) of 0.903. Applied to `r nrow(test_data)` unsigned 2026 wide receivers, the model projects a short-term market where every scored player has a modal one-year contract, the mean one-year probability is `r pct_num(xgb_summary_stats$mean_p1, 1)`, and the median expected annual value is `r money_m(xgb_summary_stats$median_expected_aav, 2)`. Player-level scenario outputs are available in the \href{https://rentosaijo.github.io/NFLxC/}{companion dashboard}.

\newpage

# Introduction and Background

NFL wide receiver contracts are difficult to compare because the headline value of a deal hides several distinct market decisions. A one-year prove-it contract, a two-year role-player contract, and a four-year star extension can all belong to the same position group, yet they imply different levels of team commitment and player leverage. For readers less familiar with football, a wide receiver is an offensive player whose primary role is to catch passes. Contract term is the reported number of contract years, and average annual value, or AAV, is total reported value divided by term. Because the NFL salary cap changes substantially over time, this project models AAV as a share of the league cap at signing rather than as raw dollars. An annual salary of \(\$10\) million carries different meaning under a league cap of \(\$120\) million than under a league cap of \(\$300\) million, so cap share gives a cleaner cross-era target. Rookie and rookie-adjacent deals are excluded because they are governed by a different pricing mechanism; keeping them in the same sample would mix the veteran free-agent market with the rookie wage scale.

The project is organized around three research questions. Can the term of a veteran wide receiver contract be predicted from public information available before signing? Can the contract's AAV as a percentage of the salary cap be predicted with useful accuracy? Which contract-history, career-stage, production, team-context, and market variables explain those predictions? These questions are significant because public contract projections are usually presented as single dollar values, while actual negotiations are closer to menus of possible terms and prices. This motivates the two-model architecture used throughout the project. The term model estimates probabilities over one, two, three, four, and five-or-more-year outcomes, and the compensation model estimates cap-share AAV under each supplied term scenario. The final output is therefore a probability-weighted set of contract paths rather than only a ranked list of dollar forecasts.

# Data and Methods

The contract outcome data come from Spotrac wide receiver contract exports grouped by decade [@spotrac]. Those exports provide player names, teams, signing ages, start and end years, reported term, total value, average salary, cap share at signing, and guarantee-related fields. The contract records are enriched with nflverse player references, rosters, combine measurements, regular receiving statistics, snap counts, Next Gen receiving metrics, team offensive context, and schedules through `nflreadr` [@nflreadr]. ESPN transaction data are used to recover signing dates when possible [@espntransactions] so each contract uses only information available before signing; unmatched dates use a March 15 offseason proxy that is recorded in the data. After cleaning, the modeling universe contains `r comma_num(nrow(train_data))` signed veteran contracts from `r min(train_data$startYear)`-`r max(train_data$startYear)`, `r nrow(validate_data)` signed 2026 validation contracts, and `r nrow(test_data)` unsigned 2026 free-agent candidates.

Cleaning is designed so that each row represents one veteran contract and each predictor describes the player before signing. The workflow standardizes names, parses contract economics, converts AAV percent to decimal cap share, repairs implausible end years, removes yearly rows already covered by an earlier multi-year deal, and builds prior-contract features. Player matching uses normalized names, age, rookie-season plausibility, roster history, and position context. Recent performance is based on the two most recent fully completed seasons before signing, with the most recent season weighted twice as heavily and missingness tracked with availability indicators rather than treated as a real zero.

The modeling recipe is implemented with tidymodels [@tidymodels]. Identifiers, audit fields, signing team, and current-contract information are excluded because they would either leak the answer or fail to generalize to unsigned players. Preprocessing handles unknown and new categorical levels, imputes missing predictors, removes zero-variance terms, and normalizes numeric inputs. The term model predicts one, two, three, four, or five-plus years, with longer contracts grouped together because they are rare. The compensation model predicts cap-share AAV under a supplied term scenario. Model selection begins with an untuned baseline screen. The predictive side covers random forest, XGBoost, LightGBM, and neural networks. The interpretable side covers multinomial logistic regression for term, ordinary linear regression for AAV%, ridge, lasso, elastic net, decision trees, and null models. Table \@ref(tab:baseline-screen) summarizes the best baseline result in each task-by-model-role lane. We then fine-tune XGBoost for both predictive tasks, lasso multinomial regression for term, and ridge regression for AAV%. Model selection uses 2002-2025 contracts for training and signed 2026 contracts for validation, with five grouped folds by player used during tuning so the same player's contract history does not leak across folds.

```{r baseline-screen-table, results = 'asis'}
baseline_screen_tbl |>
  table_latex(
    col_names = c('Task', 'Model Role', 'Best Baseline', 'Validation Evidence'),
    caption = '(\\#tab:baseline-screen) Best untuned baseline results by task and model role.',
    align = c('l', 'l', 'l', 'l'),
    wrap_col = 4,
    wrap_width = '2.35in',
    font_size = 6.5
  )
```

# Results and Discussion

Figure \@ref(fig:market-scatter) shows the main forecast pattern. The term task has to be evaluated carefully because the signed 2026 validation set is heavily concentrated in one-year contracts. A model can look accurate by predicting one year for nearly everyone, but that does not mean its probabilities are calibrated or useful for comparing players. For that reason, the selected term model is judged primarily with log loss and Brier score, which reward better probability estimates. After tuning, the selected XGBoost term model improves to log loss 0.537 and Brier 0.057, while the selected XGBoost compensation model explains most held-out variation in cap-share AAV. Its RMSE of 0.0057 represents about 0.57 percentage points of the cap, or roughly \(\$1.7\) million at the implied 2026 cap. The first research question therefore receives a qualified answer: contract term can be predicted, but a probability distribution over terms is more useful than a single predicted class. The second research question has a clearer answer: cap-share AAV is meaningfully predictable once the model conditions on term.

The unsigned 2026 wide receiver group is projected to be overwhelmingly short-term. All `r nrow(summary_xgb)` scored players have one year as their modal term, and the average one-year probability is `r pct_num(xgb_summary_stats$mean_p1, 1)`. Price is less compressed than term. The mean expected AAV is `r money_m(xgb_summary_stats$mean_expected_aav, 2)`, the median is `r money_m(xgb_summary_stats$median_expected_aav, 2)`, and the maximum is `r money_m(xgb_summary_stats$max_expected_aav, 2)`, so the projected market is right-skewed. Deebo Samuel, Jauan Jennings, and Keenan Allen rise to the top because they combine stronger recent or prior production signals with nontrivial probabilities of multi-year outcomes. Even for those players, however, the modal forecast remains one year. Their expected values should be read as probability-weighted scenario summaries rather than direct predictions of long contracts. Table \@ref(tab:compact-projections) lists the highest expected-AAV cases and shows how much probability remains on two-or-more-year outcomes.

```{r market-scatter-plot, fig.width = 6.3, fig.height = 2.2, fig.cap = '(\\#fig:market-scatter) One-year probability versus expected AAV for unsigned 2026 wide receivers.'}
label_players <- summary_xgb |>
  arrange(desc(expectedAavMillions)) |>
  slice_head(n = 7)

summary_xgb |>
  mutate(top_label = if_else(contractRowId %in% label_players$contractRowId, playerName, NA_character_)) |>
  ggplot(aes(x = .pred_1, y = expectedAavMillions)) +
  geom_point(aes(size = expectedTerm), color = '#2F6CA8', alpha = 0.68) +
  geom_text(aes(label = top_label), hjust = -0.05, vjust = 0.25, size = 2.15, check_overlap = TRUE) +
  scale_x_continuous(labels = percent_format(accuracy = 1), limits = c(0.5, 1.0)) +
  scale_y_continuous(labels = dollar_format(prefix = '$', suffix = 'M')) +
  scale_size_continuous(range = c(1.4, 4.5), guide = 'none') +
  labs(x = 'Predicted Probability of a One-Year Contract', y = 'Expected AAV') +
  theme(axis.title = element_text(size = 8), axis.text = element_text(size = 7))
```

```{r compact-projections-table, results = 'asis'}
compact_projection_tbl |>
  table_latex(
    col_names = c('Player', 'Expected Term', 'P(1 Year)', 'P(2+ Years)', 'Expected AAV'),
    caption = '(\\#tab:compact-projections) Highest expected AAV projections from the predictive model pair.',
    align = c('l', 'r', 'r', 'r', 'r'),
    font_size = 6.5
  )
```

The third research question concerns what information is built into the forecasts. The predictive and interpretable models point to the same broad distinction. Term is driven most strongly by contract history and career stage. The boosted term model leans on observed contract number, years of experience, player-match quality, previous contract timing, age, and previous market value. In the lasso audit model, one-year scores are shaped by age, previous-contract timing, and recent game availability; longer-term classes retain production, team-context, and continuity signals. That means the model treats term as a market-commitment decision first and a pure receiving-production decision second. Compensation is tied more directly to receiving value. Weighted receiving yards, receiving yards per game, first downs per game, target share, prior cap share, and the supplied term scenario dominate the boosted importance table. In the ridge audit model, the largest coefficients are term indicators and previous-contract market variables, followed by receiving and snap-context features. The two outcomes are related but not interchangeable; term and price are predictable from different kinds of information.

Player-level term-price menus are available in the \href{https://rentosaijo.github.io/NFLxC/}{companion dashboard}.

# Limitations and Future Work

Limitations mostly come from measurement and private context. Proxy signing dates can blur the pre-signing window, and name/roster matching is weakest for fringe players. Sparse pre-2020 route data make target volume hard to separate from route skill; route participation, separation, target depth/quality, and injury availability would sharpen that split. Public contracts omit guarantees, incentives, option years, medical risk, cap room, depth-chart need, and agent leverage, so AAV is incomplete. Future work should automate Spotrac refreshes, model guarantees/incentives, add interval and calibration checks, and score cap-and-need team scenarios.

\newpage

# Appendix

## Appendix A: Data Construction

The full cleaning workflow begins with Spotrac contract exports, keeps wide receiver rows, standardizes names and teams, parses contract economics, converts cap share to a decimal, buckets five-plus-year contracts, removes entry-like deals, and removes yearly rows already contained inside a prior multi-year contract. The contract spine is enriched with player-reference, roster, combine, regular-stat, snap-count, Next Gen receiving, team, and schedule data. Transaction dates are matched from ESPN by player, team, and contract clues when possible; otherwise the workflow uses the offseason proxy described in the main paper. Tables \@ref(tab:data-splits) and \@ref(tab:coverage) summarize the cleaned data splits and feature coverage.

```{r data-splits-table, results = 'asis'}
split_tbl |>
  mutate(rows = comma_num(rows), players = comma_num(players)) |>
  table_latex(
    col_names = c('Split', 'Role', 'Rows', 'Players', 'Start Years'),
    caption = '(\\#tab:data-splits) Modeling data after cleaning.',
    align = c('l', 'l', 'r', 'r', 'c'),
    wrap_col = 2,
    wrap_width = '2.7in'
  )
```

```{r coverage-table, results = 'asis'}
coverage_tbl |>
  mutate(train_coverage = pct_num(train_coverage, 1)) |>
  table_latex(
    col_names = c('Feature Family', 'Training Coverage'),
    caption = '(\\#tab:coverage) Training coverage for major external feature families.',
    align = c('l', 'r')
  )
```

Figure \@ref(fig:term-dist) explains why the paper emphasizes probability quality. One-year contracts are `r pct_num(mean(train_data$term == 1), 1)` of the training data and `r pct_num(mean(validate_data$term == 1), 1)` of the validation data.

```{r term-distribution, fig.width = 5.8, fig.height = 3.0, fig.cap = '(\\#fig:term-dist) Term distribution in the training and validation data. Five means five or more years.'}
term_dist_tbl |>
  mutate(term = factor(term, levels = 1:5, labels = c('1', '2', '3', '4', '5+'))) |>
  ggplot(aes(x = term, y = pct, fill = split)) +
  geom_col(position = position_dodge(width = 0.75), width = 0.65) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_manual(values = c('Train' = '#2F6CA8', 'Validate' = '#C44E52')) +
  labs(x = 'Contract Term', y = 'Share of Contracts', fill = NULL) +
  theme(legend.position = 'top')
```

## Appendix B: Model Comparison and Tuning

The baseline stage compared null models, linear and penalized generalized linear models, neural networks, decision trees, random forests, XGBoost, and LightGBM where available. Tuning then focused on the strongest predictive and interpretable candidates from the baseline screen. XGBoost used Bayesian optimization; the penalized models used regularization grids. Table \@ref(tab:optimized) reports the selected tuned comparison used in the paper.

```{r optimized-results, results = 'asis'}
opt_tbl |>
  filter(model != 'Random forest') |>
  table_latex(
    col_names = c('Task', 'Model', 'Validation Metric 1', 'Validation Metric 2', 'Validation Metric 3', 'Cross-Validation Metric'),
    caption = '(\\#tab:optimized) Tuned model comparison on the signed 2026 validation set.',
    align = c('l', 'l', 'r', 'r', 'r', 'r')
  )
```

Tables \@ref(tab:baseline-term) and \@ref(tab:baseline-aav) give the full successful baseline screens for the term and compensation tasks.

```{r appendix-baseline-term, results = 'asis'}
term_baseline |>
  filter(status == 'ok') |>
  arrange(mnLogLoss) |>
  transmute(
    Model = dplyr::case_when(
      modelId == 'xgboost' ~ 'XGBoost',
      modelId == 'random_forest_ranger' ~ 'Random forest',
      modelId == 'lasso_glmnet' ~ 'Lasso',
      modelId == 'elastic_net_glmnet' ~ 'Elastic net',
      modelId == 'lightgbm' ~ 'LightGBM',
      modelId == 'ridge_glmnet' ~ 'Ridge',
      modelId == 'null_majority' ~ 'Null majority',
      modelId == 'multinomial_logit_nnet' ~ 'Multinomial logistic',
      modelId == 'mlp_nnet' ~ 'Neural net',
      modelId == 'decision_tree_rpart' ~ 'Decision tree',
      TRUE ~ modelId
    ),
    `Log Loss` = round(mnLogLoss, 3),
    `ROC AUC` = round(rocAuc, 3),
    Brier = round(brier, 3),
    Accuracy = pct_num(accuracy, 1)
  ) |>
  table_latex(
    col_names = c('Model', 'Log Loss', 'ROC AUC', 'Brier', 'Accuracy'),
    caption = '(\\#tab:baseline-term) Baseline term models ordered by validation log loss.',
    align = c('l', 'r', 'r', 'r', 'r')
  )
```

```{r appendix-baseline-aav, results = 'asis'}
aav_baseline |>
  filter(status == 'ok') |>
  arrange(rmse) |>
  transmute(
    Model = dplyr::case_when(
      modelId == 'xgboost' ~ 'XGBoost',
      modelId == 'random_forest_ranger' ~ 'Random forest',
      modelId == 'ridge_glmnet' ~ 'Ridge',
      modelId == 'lightgbm' ~ 'LightGBM',
      modelId == 'linear_lm' ~ 'Linear regression',
      modelId == 'decision_tree_rpart' ~ 'Decision tree',
      modelId == 'elastic_net_glmnet' ~ 'Elastic net',
      modelId == 'mlp_nnet' ~ 'Neural net',
      modelId == 'lasso_glmnet' ~ 'Lasso',
      modelId == 'null_mean' ~ 'Null mean',
      TRUE ~ modelId
    ),
    RMSE = round(rmse, 4),
    MAE = round(mae, 4),
    `R-squared` = round(rsq, 3)
  ) |>
  table_latex(
    col_names = c('Model', 'RMSE', 'MAE', 'R-squared'),
    caption = '(\\#tab:baseline-aav) Baseline compensation models ordered by validation RMSE.',
    align = c('l', 'r', 'r', 'r')
  )
```

Table \@ref(tab:hyperparameters) records the selected XGBoost settings used for the final predictive models.

```{r appendix-hyperparameters, results = 'asis'}
hyper_tbl <- bind_rows(
  term_opt |>
    filter(modelId == 'term_xgb') |>
    transmute(
      Model = 'Term XGBoost',
      mtry = mtry,
      trees = trees,
      min_n = min_n,
      tree_depth = tree_depth,
      learn_rate = learn_rate,
      loss_reduction = loss_reduction,
      sample_size = sample_size
    ),
  aav_opt |>
    filter(modelId == 'aavPerc_xgb') |>
    transmute(
      Model = 'Compensation XGBoost',
      mtry = mtry,
      trees = trees,
      min_n = min_n,
      tree_depth = tree_depth,
      learn_rate = learn_rate,
      loss_reduction = loss_reduction,
      sample_size = sample_size
    )
)

hyper_tbl |>
  mutate(across(where(is.numeric), ~ round(.x, 6))) |>
  table_latex(
    col_names = c('Model', 'mtry', 'Trees', 'Minimum N', 'Tree Depth', 'Learning Rate', 'Loss Reduction', 'Sample Size'),
    caption = '(\\#tab:hyperparameters) Selected XGBoost hyperparameters.',
    align = c('l', rep('r', 7)),
    font_size = 7
  )
```

## Appendix C: Interpretation Artifacts

The XGBoost models are interpreted through global gain importance. Importance is not a causal effect or a signed marginal coefficient; it reports which transformed predictors most improved split quality across the fitted trees. Figures \@ref(fig:term-importance) and \@ref(fig:aav-importance) show the predictive models' highest-gain features.

```{r term-importance-plot, fig.width = 6.1, fig.height = 3.6, fig.cap = '(\\#fig:term-importance) Top gain-importance features for the predictive term model.'}
term_xgb_imp |>
  slice_head(n = 15) |>
  mutate(Feature = fct_reorder(clean_feature(Feature), Gain)) |>
  ggplot(aes(x = Gain, y = Feature)) +
  geom_col(fill = '#2F6CA8') +
  labs(x = 'Gain Importance', y = NULL)
```

```{r aav-importance-plot, fig.width = 6.1, fig.height = 3.6, fig.cap = '(\\#fig:aav-importance) Top gain-importance features for the predictive conditional AAV% model.'}
aav_xgb_imp |>
  slice_head(n = 15) |>
  mutate(Feature = fct_reorder(clean_feature(Feature), Gain)) |>
  ggplot(aes(x = Gain, y = Feature)) +
  geom_col(fill = '#4C956C') +
  labs(x = 'Gain Importance', y = NULL)
```

Tables \@ref(tab:lasso-sparsity) and \@ref(tab:lasso-coefs) summarize the lasso audit model. The sparsity table shows how many predictors remain active by term class, while the coefficient table shows the strongest retained signals within each class.

```{r appendix-lasso-sparsity, results = 'asis'}
term_ll_coef |>
  filter(feature != '(Intercept)', coefficient != 0) |>
  count(class, name = 'non_zero_coefficients') |>
  mutate(class = if_else(class == 5, '5+', as.character(class))) |>
  table_latex(
    col_names = c('Term Class', 'Non-Zero Coefficients'),
    caption = '(\\#tab:lasso-sparsity) Class-specific sparsity in the lasso multinomial term model.',
    align = c('c', 'r')
  )
```

```{r appendix-lasso-coefficients, results = 'asis'}
term_ll_coef |>
  filter(feature != '(Intercept)', coefficient != 0) |>
  mutate(
    class = if_else(class == 5, '5+', as.character(class)),
    abs_value = abs(coefficient)
  ) |>
  arrange(class, desc(abs_value)) |>
  group_by(class) |>
  slice_head(n = 5) |>
  ungroup() |>
  transmute(
    `Term Class` = class,
    Feature = clean_feature(feature),
    Coefficient = round(coefficient, 3),
    `Absolute Value` = round(abs_value, 3)
  ) |>
  table_latex(
    col_names = c('Term Class', 'Feature', 'Coefficient', 'Absolute Value'),
    caption = '(\\#tab:lasso-coefs) Five largest absolute nonzero lasso coefficients by term class.',
    align = c('c', 'l', 'r', 'r'),
    font_size = 6.5
  )
```

Table \@ref(tab:ridge-coefs) gives the largest absolute coefficients from the ridge compensation audit model.

```{r appendix-ridge-coefficients, results = 'asis'}
aav_rl_coef |>
  filter(feature != '(Intercept)') |>
  mutate(abs_value = abs(coefficient)) |>
  arrange(desc(abs_value)) |>
  transmute(
    Feature = clean_feature(feature),
    Coefficient = round(coefficient, 6),
    `Absolute Value` = round(abs_value, 6)
  ) |>
  slice_head(n = 15) |>
  table_latex(
    col_names = c('Feature', 'Coefficient', 'Absolute Value'),
    caption = '(\\#tab:ridge-coefs) Largest absolute coefficients in the ridge compensation model.',
    align = c('l', 'r', 'r'),
    font_size = 7
  )
```

## Appendix D: Additional Forecast Outputs

Table \@ref(tab:top-projections) expands the main-paper projection table to the ten highest expected-AAV players. Figure \@ref(fig:scenario-paths) shows how predicted AAV changes when the same player is evaluated under different term scenarios, with point size representing the term probability assigned by the term model.

```{r top-projections-table, results = 'asis'}
top_projection_tbl |>
  table_latex(
    col_names = c('Player', 'Age', 'Previous Team', 'Expected Term', 'P(1 Year)', 'P(2+ Years)', 'Expected AAV', 'Modal AAV'),
    caption = '(\\#tab:top-projections) Ten highest expected AAV projections from the predictive model pair.',
    align = c('l', 'r', 'c', 'r', 'r', 'r', 'r', 'r'),
    font_size = 7
  )
```

```{r scenario-paths-plot, fig.width = 6.2, fig.height = 3.6, fig.cap = '(\\#fig:scenario-paths) Scenario-level AAV paths for the three highest expected-AAV players.'}
top_three <- summary_xgb |>
  arrange(desc(expectedAavMillions)) |>
  slice_head(n = 3) |>
  pull(playerName)

scenarios_xgb |>
  filter(playerName %in% top_three) |>
  mutate(playerName = factor(playerName, levels = top_three)) |>
  ggplot(aes(x = termScenario, y = aavPredictionMillions, group = playerName, color = playerName)) +
  geom_line(linewidth = 0.8) +
  geom_point(aes(size = termProbability), alpha = 0.85) +
  scale_x_continuous(breaks = 1:5, labels = c('1', '2', '3', '4', '5+')) +
  scale_y_continuous(labels = dollar_format(prefix = '$', suffix = 'M')) +
  scale_size_continuous(labels = percent_format(accuracy = 1), range = c(1.5, 5)) +
  labs(x = 'Term Scenario', y = 'Predicted AAV', color = NULL, size = 'Term Probability') +
  guides(
    size = guide_legend(order = 1, nrow = 1),
    color = guide_legend(order = 2, nrow = 1)
  ) +
  theme(
    legend.position = 'bottom',
    legend.box = 'vertical',
    legend.text = element_text(size = 7),
    legend.title = element_text(size = 7)
  )
```

\newpage

# References
